### 2023-02-23 doc/definitions.tex
start blueprint

\section{Introduction}

\begin{definition}
\label{local_field}
\lean{local_field}
\leanok
A \textit{(nonarchimedean) local field} is a field complete with respect to a discrete
valuation and with finite residue field.
\end{definition}

\begin{definition}
\label{mixed_char_local_field}
\lean{mixed_char_local_field}
\leanok
A \textit{mixed characteristic local field} is a finite field extension of the field 
$\mathbb{Q}_p$ of $p$-adic numbers, for some prime $p$.
\end{definition}

\begin{definition}
\label{eq_char_local_field}
\lean{eq_char_local_field}
\leanok
An \textit{equal characteristic local field} is a finite field extension of the field 
$\mathbb{F}_p((X))$, for some prime $p$.
\end{definition}

\begin{lemma}
\label{mixed_char_local_field.to_local_field}
\lean{mixed_char_local_field.to_local_field}
\uses{local_field, mixed_char_local_field}
A mixed characteristic local field is a local field.
\end{lemma}

\begin{lemma}
\label{eq_char_local_field.to_local_field}
\lean{eq_char_local_field.to_local_field}
\uses{local_field, eq_char_local_field}
An equal characteristic local field is a local field.
\end{lemma}

\begin{definition}
\label{def_local_field}
\lean{blueprint.first_test}
A local field is a field
\end{definition}

\begin{proof}
\leanok
Omitted, but it uses \cite{SerCF} and also \cite{Se2}.
\end{proof}