### 2023-04-20 doc/definitions.tex

\section{Generalities}
Fix a field $K$.
\begin{definition}
\label{def:discrete_val}
\lean{is_discrete}
\leanok
A valuation $v\colon K \to \ZZ$ on $K$ is discrete if it is non-archimedean and surjective.
\end{definition}
We have the

\begin{proposition}[{\cite[Chap I,\S1, Proposition 1]{Se2}}]
\label{prop:unit_ball_dvr}
\lean{discrete_valuation.integer.discrete_valuation_ring}
If $K$ is complete with respect to a discrete valuation $v$, then its unit ball $\unitball{K}$ is a discrete valuation ring.
\end{proposition}

\begin{lemma}[{\cite[Chap II,\S2, last part of Proposition 3]{Se2}}] 
\label{lemma:val_L}
\lean{discrete_valuation.hw}
\lean{discrete_valuation.complete_of_finite}
\lean{discrete_valuation.is_discrete_of_finite}
If $K$ is complete with respect to a discrete valuation $v$ and if $L/K$ is a finite extension, then $L$ has a discrete valuation $w\colon L \twoheadrightarrow \ZZ$ inducing $v$ and $L$ is complete with respect to $w$.
\end{lemma}

Moreover,
\begin{proposition}[{\cite[Chap II,\S2, Proposition 3]{Se2}}]
\label{prop:unitball_L}
\lean{discrete_valuation.integral_closure_eq_integer}
\lean{discrete_valuation.integral_closure.discrete_valuation_ring}
\lean{discrete_valuation.integral_closure_finrank}
If $K$ is complete with respect to a discrete valuation $v$ and if $L/K$ is a finite extension, then the integral closure of $\unitball{K}$ inside $L$ coincides $\unitball{L}$ and so, in particular, it is a discrete valuation ring by~\ref{prop:unit_ball_dvr}. Moreover, $\unitball{L}$ is a finite, free $\unitball{K}$-module of rank $n=[L:K]$.
\end{proposition}
\section{Local Fields}
\begin{definition}
\label{local_field}
\lean{local_field}
\leanok
A \textit{(nonarchimedean) local field} is a field complete with respect to a discrete
valuation and with finite residue field.
\end{definition}

\begin{definition}
\label{mixed_char_local_field}
\lean{mixed_char_local_field}
\leanok
A \textit{mixed characteristic local field} is a finite field extension of the field 
$\Qp$ of $p$-adic numbers, for some prime $p$.
\end{definition}

\begin{definition}
\label{eq_char_local_field}
\lean{eq_char_local_field}
\leanok
An \textit{equal characteristic local field} is a finite field extension of the field 
$\laurentseries$, for some prime $p$.
\end{definition}

\begin{lemma}
\label{mixed_char_local_field.local_field}
\lean{mixed_char_local_field.local_field}
\uses{local_field, mixed_char_local_field}
A mixed characteristic local field is a local field.
\end{lemma}

\begin{lemma}
\label{eq_char_local_field.local_field}
\lean{eq_char_local_field.local_field}
\uses{local_field, eq_char_local_field}
An equal characteristic local field is a local field.
\end{lemma}
\subsection{Ramification Index}
\section{Global to Local}
\begin{proposition}
\label{prop:number_field_localization}
\lean{number_field.adic_completion.mixed_char_local_field}
Let $F$ be a number field and let $v$ be a finite place of residue characteristic $p$. Then $F_v$ is a mixed characteristic local field with residue characteristic $p$.
\end{proposition}

\begin{proposition}
\label{prop:function_field_localization}
\lean{function_field.adic_completion.eq_char_local_field}
Let $F$ be a function field and let $v$ be any place. Then $F_v$ is an equal characteristic local field with residue characteristic $p$.
\end{proposition}