### 2023-05-25 blueprint/src/web.tex
changes to blueprint

\documentclass{report}

\usepackage{amsmath, amsthm}
\usepackage{graphicx}
\DeclareGraphicsExtensions{.svg,.png,.jpg}
\usepackage[capitalize]{cleveref}
\usepackage[showmore, dep_graph, coverage, project=../../]{blueprint}

\usepackage{tikz-cd}
\usepackage{tikz}
%
%\input{macros_common}
%\input{macros_web}

\home{https://github.com/mariainesdff/local_class_field_theory}
\github{https://github.com/mariainesdff/local_class_field_theory}

\title{The sphere eversion project}
\author{Patrick Massot \and Oliver Nash \and Floris van Doorn}

\begin{document}
\maketitle
\input{content}
\end{document}