### 2023-06-16 CPP_paper/paper.tex
improved instances in basic

\documentclass[sigplan,10pt,anonymous,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
\include{preamble}

\author[M.~I.~de Frutos Fernández]{María Inés de Frutos Fernández}
\author[F.~A.~E.~Nuccio]{Filippo Alberto Edoardo Nuccio Mortarino Majno di Capriglio}
\title{later}
\begin{document}
\maketitle
\section{Introduction}
\section {Discrete Valuation Rings}
\subsection{Valuations}
\begin{itemize}
\item Properties of valuations valued in $\ZZ_{m0}$, and the \texttt{multiplicative} and \texttt{with\_zero} classes, and clashes with "usual" intuition. Keep in mind that \texttt{with\_zero multiplicative} $\ZZ$ is "the same" as $q^{-\ZZ}\cup \{0\}$.
\item Difference between \texttt{valued} (which is a class, and carries a topology/uniformity) and \texttt{valuation} that is just a function.
\item We put \texttt{valued} instances on fields but "never" on rings that are not fields.
\item There is something called \texttt{valuation\_ring} that has (almost) nothing to do; but \texttt{valuation\_subring} has.
\end{itemize}
\subsection{Definition of Discrete Valuation Rings}
\begin{itemize}
\item True Math definition and the \texttt{mathlib\_tfae} (in particular, local and PID)
\item \texttt{is\_discrete} : observe that if the value group were $2\ZZ$ this would still be mathematically "discrete", but we make the choice of surjectivity in coherence with Serre/so that the notion of equivalence collapses to equality: there is a unique normalized one.
\item \texttt{uniformizer}'s and equivalence of their existence with discreteness
\item The ring of integers of a field with a discrete valuation (be it valued or not) is a DVR
\end{itemize}
\subsection{Extensions of Complete Discrete Valuation Rings}
\begin{itemize}
\item 
\end{itemize}
\section{Local Fields}
\subsection{Equal characteristic}
insert here the iso between Laurent series and the completion
\subsection{Mixed characteristic}
\section{Discussion}
\begin{itemize}
\item Future work (not too much, but towards LCFT)
\item Related works
\end{itemize}
\subsection{Implementation Details}
\begin{itemize}
\item insert here the "double" $Q_p$.
\item 
\end{itemize}
%\section*{Acknowledgement}--LATER (fundings?)
\end{document}