### 2023-06-16 CPP_paper/paper.tex
modifying extension file

\documentclass[sigplan,10pt,anonymous,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
\include{preamble}

\author[M.~I.~de Frutos Fernández]{María Inés de Frutos Fernández}
\author[F.~A.~E.~Nuccio]{Filippo Alberto Edoardo Nuccio Mortarino Majno di Capriglio}
\title{later}
\begin{document}
\maketitle
\section{Introduction}
\section {Discrete Valuation Rings}
\subsection{Valuations}
\begin{itemize}
\item Properties of valuations valued in $\ZZ_{m0}$, and the \texttt{multiplicative} and \texttt{with\_zero} classes, and clashes with "usual" intuition. Keep in mind that \texttt{with\_zero multiplicative} $\ZZ$ is "the same" as $q^{-\ZZ}\cup \{0\}$.
\item Difference between \texttt{valued} (which is a class, and carries a topology/uniformity) and \texttt{valuation} that is just a function.
\item We put \texttt{valued} instances on fields but "never" on rings that are not fields.
\item There is something called \texttt{valuation\_ring} that has (almost) nothing to do; but \texttt{valuation\_subring} has.
\item Comment about the choice of \texttt{.integers} vs. \texttt{.valuation\_subring}
\end{itemize}
\subsection{Definition of Discrete Valuation Rings}
\begin{itemize}
\item True Math definition and the \texttt{mathlib\_tfae} (in particular, local and PID)
\item \texttt{is\_discrete} : observe that if the value group were $2\ZZ$ this would still be mathematically "discrete", but we make the choice of surjectivity in coherence with Serre/so that the notion of equivalence collapses to equality: there is a unique normalized one.
\item \texttt{uniformizer}'s and equivalence of their existence with discreteness
\item The ring of integers of a field with a discrete valuation (be it valued or not) is a DVR; and the adic valuation (ref to María Inés' paper) on the field of fractions of a DVR \texttt{is\_discrete}.
\end{itemize}
\subsection{Extensions of Complete Discrete Valuation Rings}
\begin{itemize}
\item Recall (María Inés' paper): there is a norm associated to a valuation; and if the field below is complete and the extension is finite, then this norm is unique (extending the one below).
\item Speak about the work to define the normalized \texttt{discrete} valuation "above"
\item [MAYBE] Mention the lemma \texttt{w\_le\_one\_iff\_disc\_norm\_extension\_le\_one}
\item The proof (from previous project) and instance that a finite extension of a complete field is again complete
\item The lemma \path{integral_closure_eq_integer} and the theorem \path{dvr_of_finite_extension}
\end{itemize}
\section{Local Fields}
\subsection{Equal characteristic}
insert here the iso between Laurent series and the completion
\subsection{Mixed characteristic}
\section{Discussion}
\begin{itemize}
\item Future work (not too much, but towards LCFT)
\item Related works
\end{itemize}
\subsection{Implementation Details}
\begin{itemize}
\item bundled vs unbundled \texttt{uniformizer}'s and the choice of having them in K or in R or in R.int\dots
\item insert here the "double" $Q_p$.
\item the loop forcing the instance \texttt{discretely\_normed\_field} to be only a local one
\item [MAY BE] Discuss working in $\ZZ_{m0}$ vs in $\RR$ for the DVR\_extension
\end{itemize}
%\section*{Acknowledgement}--LATER (fundings?)
\end{document}